### 2024-02-16 paper_sisc_first_revision/localpsf.tex
ricker vs frog ellipsoids figure added
added discussion of what happens when there are negative numbers in the integral kernel

We focus on approximating operators that satisfy the following conditions: 
\begin{enumerate}
	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well-defined for all $x\in \Omega$.
	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
	\item The integral kernel is non-negative\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness. The operator $\Aop$ need not be positive semi-definite to use the proposed method, and positive semi-definite operators need not have a non-negative integral kernel.} in the sense that $$\Aker(y,x) \ge 0 \quad \text{for all} \quad (y,x) \in \Omega \times \Omega.$$
\end{enumerate}
The proposed method may still perform well if these conditions are relaxed slightly. It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region. The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$ so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and so long as the negative numbers are comparatively small. If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.

@@

\section*{Acknowledgments}
We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. The seismic Hessian impulse response image was provided to us by Mathew Hu. We thank Trevor Heise for editing suggestions. We thank Georg Stadler for help with the domain setup for the ice sheet problem.

>>>>

We focus on approximating operators that satisfy the following conditions: 
\begin{enumerate}
	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well-defined for all $x\in \Omega$.
	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
	\item The integral kernel is non-negative\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness. The operator $\Aop$ need not be positive semi-definite to use the proposed method, and positive semi-definite operators need not have a non-negative integral kernel.} in the sense that $$\Aker(y,x) \ge 0 \quad \text{for all} \quad (y,x) \in \Omega \times \Omega.$$
\end{enumerate}
The proposed method may still perform well if these conditions are relaxed slightly.

%If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.

It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region. In this case the method can still be applied, but the ellipsoid impulse response support estimates will be large. Therefore, only a small number of impulse responses can be computed per batch. This means that more batches, and hence more operator applies, are needed to form an accurate approximation.

%The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$ (Assumption 3) so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and so long as the negative numbers are comparatively small.

If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the PSF method. In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3. We study two integral kernel examples, both of which are parameterized by a quantity that controls how negative the kernels are. We make the following observations:
\begin{itemize}
	\item The larger and more numerous the negative numbers are, the more inaccurate the ellipsoid support estimate is.
	\item The further away from the center of the ellipsoid the negative numbers are, the more influence they have on the quality of the ellipsoid support estimate. This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
	\item Negative numbers in the integral kernel affect the method more if they are isolated, and less if they are balanced by nearby positive numbers.
	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries. After that threshold is crossed, the estimation rapidly transitions to performing poorly and ultimately failing.
\end{itemize} 
%The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the method fails. For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles. In the bottom two rows, we see that the method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.

%Location and magnitude of negatives matter. Larger negatives: bad, smaller negatives good.  Negatives closer to center: good, further from center: bad. Correlated negatives: bad, randomly distributed: good. Good Good Good Fail quick
%
%Non-local: will work, but choose few ellipsoids, so inefficient

\begin{figure}
	\centering
	{
		%	\setlength{\tabcolsep}{0.1em}
		%	\renewcommand{\arraystretch}{2.0}
		\begin{tabular}{cccc}
	\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=1.0.png} & 
	\includegraphics[scale=0.3, valign=c]{frog_1d_a=1.0.png} & 
	\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.0.png} & 
	\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.0.png} \\
	%
	\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=20.0.png} & 
	\includegraphics[scale=0.3, valign=c]{frog_1d_a=20.0.png} & 
	\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.23.png} & 
	\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.23.png} \\
	%
	\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=27.0.png} & 
	\includegraphics[scale=0.3, valign=c]{frog_1d_a=27.0.png} & 
	\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.249.png} & 
	\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.249.png} 
	\end{tabular}
	}
	\caption{Illustration of the influence of negative numbers in the integral kernel on the robustness of the ellipsoid estimates for the supports of impulse responses. Left two columns: spatially variant blurring kernel given in Equation \ref{eq:frog_kernel}. Right two columns: Ricker wavelet-type kernel given by $\Aker(y,x) = (1 - a \theta)\exp\left(-\theta/2\right)$, where $\theta=(y-x)^T \Sigma^{-1} (y-x)$, and $\Sigma=\operatorname{diag}\left(0.0025, 0.01\right)$. Ordered from top to bottom, the results are obtained with $a\in\{1.0, 20.0, 27.0\}$ for the blur kernel, and $a\in \{0.0, 0.23, 0.249\}$ for the Ricker wavelet-type kernel. Columns 1 and 3: impulse responses for the blur and Ricker wavelet-type kernels, respectively, with estimated support ellipsoids indicated by the black ellipses. Red and blue represent positive and negative numbers in the integral kernel, respectively. Columns 2 and 4: one-dimensional slice along the horizontal line indicated in the two-dimensional plots. The dashed gray line is at zero.
	}
	\label{fig:robustness_to_negatives}
\end{figure}

@@

\section*{Acknowledgments}
We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. We thank Nicole Aretz and Trevor Heise for editing suggestions. We thank Georg Stadler for help with the domain setup for the ice sheet problem.
